The five pillars of computational reproducibility: Bioinformatics and beyond

### 2023-05-30 manuscript/5pillars.Rmd
MyST and JupyterLite

Around the same time, Project Jupyter was developed to provide a "notebook" type interface, incorporating
R, Python, Julia and other computer language chunks into documents @Perkel2018-rz.
Similarly, the [Streamlit](https://streamlit.io/) tool, which was originally invented for creating
Python-based web dashboards but is equally capable of generating data science reports
[e.g.: @Nantasenamat2023-wr].

@@

We are also watching with interest the early development of WebAssembly (Wasm) based approaches for
compute environment control.
The ability to reproduce operating systems, programming languages and workflows in the browser
opens up intriguing possibilities for more widespread reproduction and auditing without the need to
install any software at all.
For example, it is now possible to embed a Wasm-enabled R terminal inside a Quarto HTML document, which
enables in-browser reproducibility of R workflows @Balamuta2023-ma.

>>>>

Around the same time, Project Jupyter was developed to provide a "notebook" type interface, incorporating
R, Python, Julia and other computer language chunks into documents @Perkel2018-rz.
The MyST{NB} (short for Markedly Structured Text notebook) is built on top of Jupyter and has some
notable advantages around flexibility of content that make it an attractive choice for an executable
paper authoring [@DuPre2022-bk].

Similarly, the [Streamlit](https://streamlit.io/) tool, which was originally invented for creating
Python-based web dashboards but is equally capable of generating data science reports
[e.g.: @Nantasenamat2023-wr].

@@

We are also watching with interest the early development of WebAssembly (Wasm) based approaches for
compute environment control.
The ability to reproduce operating systems, programming languages and workflows in the browser
opens up intriguing possibilities for more widespread reproduction and auditing without the need to
install any software at all.
Two notable examples of this include JupyterLite, a version of JupyterLab in early development that
runs entirely in the browser @Tuloup2021-le, and WebR-enabled interactive code blocks in Quarto HTML
documents @Balamuta2023-ma.
